---
title: "Hw1_Anna_Shaljyan_Marketing_Analytics"
author: "Anna Shaljyan"
date: "9/26/2023"
output:
  html_document:
    css: custom-style.css
  pdf_document: default
  word_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include = FALSE}
#Reading the libraries
library(ggplot2)
library(ggpubr)
library(knitr)
library(diffusion)
library(openxlsx)
library(dplyr)
library(RColorBrewer)

#Path to the Excel file
excel_file <- "statistic_id236131_wholesale-sales--consumer-segment--of-treadmills-in-the-us-2007-2022.xlsx"
```

### Data Manipulation

##### I read the excel file, the data sheet that contains consumer segment wholesale sales of treadmills for 16 years.

```{r}

# Reading the second sheet ("Data") of the Excel file
df <- read.xlsx(excel_file, sheet = "Data")

initial_data <- df[2:17, ] #2007-2022
data <- df[8:14, ] #2013-2019

colnames(initial_data) <- c("Year", "Sales of treadmills")
colnames(data) <- c("Year", "Sales of treadmills")

initial_data$Year <- as.integer(initial_data$Year)
initial_data$`Sales of treadmills` <- as.integer(initial_data$`Sales of treadmills`)

data$Year <- as.integer(data$Year)
data$`Sales of treadmills` <- as.integer(data$`Sales of treadmills`)

# View the head of the data
print(initial_data)
print(data)
```

##### Now after modifying data to contain Year and Sales of treadmills we can plot it to have better understanding of the data.

```{r, fig.width=6, fig.height=4}

# Defining a color palette from RColorBrewer
year_colors <- brewer.pal(7, "Paired")  # 7 colors for 7 years

ggplot(data = data, aes(x = Year, y = `Sales of treadmills`, fill = factor(Year))) +
  geom_bar(stat = 'identity') +
  labs(title = 'Treadmill wholesale sales in U.S, in million dollars', 
       x = ' ', y = 'sales') +
  scale_x_continuous(breaks = 2013:2019) +
  scale_fill_manual(values = year_colors) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "none")  

```

#### Let's define f(t) and F(t) of bass model:
##### f (t) is the fraction of the total market that adopts at time t
##### F(t) is the fraction of the total market that has adopted up to and including time t.
##### Here p and q are innovation and immitation rates, respectively.

```{r}
bass.f <- function(t,p,q){((p+q)^2/p)*exp(-(p+q)*t)/(1+(q/p)*exp(-(p+q)*t))^2}

bass.F <- function(t,p,q){(1-exp(-(p+q)*t))/(1+(q/p)*exp(-(p+q)*t))}
```

##### We can experiment with p and q values. We can try two methods of parameter estimation: nls (non-linear least squares) and diffusion method. 

##### Important Note: as in initial sales data of 16 years there were periods with drastic sales falls or sales increases, NLS and diffusion methods where either giving error or estimating inadequately. Because of that I tried to slice the data for smaller range of year to have better and accurate estimations of p,q and m (market potential).

```{r}
sales_test = initial_data$`Sales of treadmills`
t_testing = 1:length(sales_test)

sales = data$`Sales of treadmills`
t = 1:length(sales)

print(sales_test)
print(sales)
```
#### As you can see it gives error, because of drastic 200-500 change in values for the last 3 years. Uncomment to see error. 

```{r}
# Fit the Bass diffusion model

#bass_m_test <- nls(sales_test ~ m*(((p+q)^2/p)*exp(-(p+q)*t_testing))/
#(1+(q/p)*exp(-(p+q)*t_testing))^2, start=c(list(m=sum(sales_test), 
#p=0.002, q=0.045)))

# Display the summary

#summary(bass_m_test)
```

#### NLS method of parameter estimation (p,q,m)
```{r}
# Fit the Bass diffusion model
bass_m <- nls(sales ~ m*(((p+q)^2/p)*exp(-(p+q)*t))/(1+(q/p)*exp(-(p+q)*t))^2,
              start=c(list(m=sum(sales), p=0.002, q=0.045)))

# Display the summary
summary(bass_m)
```
#### Diffusion method of parameter estimation (p,q,m)
```{r}
sales = data$`Sales of treadmills`
diff_m = diffusion(sales)
p=round(diff_m$w,4)[1]
q=round(diff_m$w,4)[2]
m=round(diff_m$w,4)[3]
diff_m

```
```{r}
# Getting the model summary
bass_summary <- summary(bass_m)

# Extracting the coefficients (including p and q)
coefficients <- coef(bass_m)

# Extracting the estimated p and q values
p_estimate <- coefficients["p"]
q_estimate <- coefficients["q"]
m_estimate <- coefficients["m"]

# Printing the estimated p and q values
cat("Estimated p:", p_estimate, "\n")
cat("Estimated q:", q_estimate, "\n")
cat("Estimated m:", m_estimate, "\n")
```
#### Plotting f(t) and histogram of treadmill sales using NLS parameter estimation
```{r, fig.width=6, fig.height=4}

# Defining ColorBrewer palette
color_palette <- "RdPu"

# Plot 1: time_ad
time_ad <- ggplot(data.frame(t = c(1:7)), aes(t)) +
  stat_function(fun = bass.f, args = c(p = 0.02208142, q = 0.04511991),
                aes(color = "Bass Model")) +
  labs(title = 'f(t)') +
  theme_minimal() +
  theme(legend.position = "none") +  
  scale_color_manual(values = c("Bass Model" = "black"))  

# Plot 2: treadmill_sales
treadmill_sales <- ggplot(data = data, aes(x = Year, y = `Sales of treadmills`,
                                           fill = factor(Year))) +
  geom_bar(stat = 'identity') +
  labs(x = " ", y = "sales") +
  scale_fill_brewer(palette = color_palette) +  
  scale_x_continuous(breaks = 2013:2019) +
  theme(legend.position = "none")  

# Arranging the plots using ggarrange
ggarrange(time_ad, treadmill_sales)


```

#### Plotting f(t) and histogram of treadmill sales using difussion parameter estimation

```{r, fig.width=6, fig.height=4}

# Defining a custom palette
my_palette <- colorRampPalette(brewer.pal(9, "PuRd"))(7)

# Plot 1: time_ad (Black)
time_ad <- ggplot(data.frame(t = c(1:7)), aes(t)) +
stat_function(fun = bass.f, args = c(p = 0.0275, q = 0.0509), 
              color = "black") +
  labs(title = 'f(t)') +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_color_identity()  

# Plot 2: treadmill_sales (PuRd palette)
treadmill_sales <- ggplot(data = data, aes(x = Year, 
                                           y = `Sales of treadmills`)) +
  geom_bar(stat = 'identity', fill = my_palette) +  
  scale_x_continuous(breaks = 2013:2019) +
  labs(x = " ", y = "Sales")

# Arranging the two plots
ggarrange(time_ad, treadmill_sales)

```

#### Estimating the number of adopters by period with NLS parameter estimation p, q and m.

```{r, fig.width=6, fig.height=4}

data$pred_sales <- bass.f(1:7, p = 0.02208142, q = 0.04511991) * 42644.88

# Defining the ColorBrewer palette
my_palette <- brewer.pal(7, "YlGn")

# Create the ggplot visualization
ggplot(data = data, aes(x = Year, y = `Sales of treadmills`)) +
  geom_bar(stat = 'identity', fill = my_palette) +  
  geom_point(mapping = aes(x = Year, y = pred_sales), color = 'red') +
  labs(x = ' ', y = 'Sales in million dollars ($)') +
  scale_x_continuous(breaks = 2013:2019, labels = 2013:2019) +
  scale_fill_identity()  
```

#### Estimating the number of adopters by period with diffusion parameter estimation p, q and m.

```{r, fig.width=6, fig.height=4}

data$pred_sales <- bass.f(1:7, p = 0.0275, q = 0.0509) * 34669.9697

# Create the ggplot
ggplot(data = data, aes(x = Year, y = `Sales of treadmills`)) +
  geom_bar(stat = 'identity', fill = brewer.pal(7, "YlGnBu")) +  
  geom_point(mapping = aes(x = Year, y = pred_sales), color = 'red') +
  labs(x = ' ', y = 'Sales in million dollars ($)') +
  scale_x_continuous(breaks = 2013:2019, labels = 2013:2019) +
  scale_fill_identity() 


```



